---
title: "Machine Learning Project"
author: "Jerry Hsieh"
date: "2014/11/20"
output: html_document
---


#### Introduction
This project is to use "Weight Lifting Exercise Dataset" and Machine Learning prediction model to predict 20 test case (tessting dataset) base on training dataset (including 19,622 samples)


#### About Data
The dataset is based on "Weight Lifting Exercise Dataset" cited below

Cite:

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz3UWkr867x

Main purpose of this dataset is to quantify "how well they do it".

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: 
exactly according to the specification (Class A) 

throwing the elbows to the front (Class B) 

lifting the dumbbell only halfway (Class C) 

lowering the dumbbell only halfway (Class D)

throwing the hips to the front (Class E)

The training dataset provides 19,622 samples with 159 variables with 5 types of activity to detect 1 correct method and 4 common mistakes (5 class total)

#### Data Preprocess

Read training dataset and testing dataset
```{r}
pml <- read.csv("pml-training.csv")
pmlTest <- read.csv("pml-testing.csv")
```

Remove first 7 columns which are mainly information purpose, not for predictors

```{r}
pml <- pml[,-seq(1:7)]
pmlTest <- pmlTest[,-seq(1:7)]
```


Remove those columns with NA's
```{r}
pml <- pml[,colSums(is.na(pml))==0]
pmlTest <- pmlTest[,colSums(is.na(pmlTest))==0]
```

For training dataset, remove columns not numeric but keep classe column.
For testing dataset, remove columns not numeric and remove problem id.
```{r}
nums1 <- sapply(pml, is.numeric)
nums1["classe"] <- TRUE
pml <- pml[nums1]
nums2 <- sapply(pmlTest, is.numeric)
nums2["problem_id"] <- FALSE
pmlTest <- pmlTest[nums2]
```

Now columns that we are using for predictors
```{r}
names(pml)
```

#### Build the model

Create training and testing dataset for cross validation inside original training dataset

```{r}
library(caret)
inTrain <- createDataPartition(pml$classe, p=0.75, list=FALSE)
pmlTraining <- pml[inTrain,]
pmlTesting <- pml[-inTrain,]
```

Create Model using "Random Forest" method, this method used in "Qualitative Activity Recognition of Weight Lifting Exercises" as well

```{r}
train_control <- trainControl(method = "cv", number = 5, preProcOptions="pca")
modFit <- train(classe ~ ., data=pmlTraining, method="rf", trControl=train_control, ntree=50)
modFit
```

Check model with testing set in training dataset
```{r}
pred <- predict(modFit, pmlTesting)
confusionMatrix(pmlTesting$classe, pred)
```

#### Apply to test dataset

Finally apply model to test dataset to predict 20 test cases
```{r}
testPred <- predict(modFit, pmlTest)
testPred
```

